### 2018-05-22 Documentation/ogtablesandtest.Rnw

\documentclass{article}
\usepackage{graphicx}
\usepackage{mathptmx}      
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{cite}
\usepackage{caption}
\usepackage{natbib}
\usepackage{longtable}
\usepackage{array}
\usepackage[table]{xcolor}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\begin{document}

% Original Tables and Figures

% Table 10

% \begin{table}[h!]
% 	\centering
% 	\caption{The Three Form PM Design}
% 	\label{tab:table10}
% 	\setlength{\tabcolsep}{1cm}
% 	\begin{tabular}{c|cccc}
% 		\toprule
% 		& \multicolumn{4}{c}{Block} \\
% 		Form & X & A & B & C \\
% 		\midrule
% 		1 & 1 & 1 & 1 & 0 \\
% 		2 & 1 & 1 & 0 & 1 \\
% 		3 & 1 & 0 & 1 & 1 \\
% 		\bottomrule
% 	\end{tabular}
% \end{table}

% Table 11

% \begin{table}[!h]
% 	\centering
% 	\caption{The Wave Missingness PM Design}
% 	\label{tab:table11}
% 	\setlength{\tabcolsep}{0.75cm}
% 	\begin{tabular}{c|ccccc}
% 		\toprule
% 		& \multicolumn{5}{c}{Month} \\
% 		Participant & 1 & 2 & 3 & 4 & 5 \\
% 		\midrule
% 		1 & 1 & 0 & 1 & 0 & 1 \\
% 		2 & 1 & 1 & 0 & 1 & 0 \\
% 		3 & 0 & 0 & 1 & 1 & 1 \\
% 		4 & 1 & 1 & 1 & 0 & 0 \\
% 		\bottomrule
% 	\end{tabular}
% \end{table}

Tables for paper

<<primary, echo=FALSE, message=FALSE, warning=FALSE, fig.pos = "H">>=
library(kableExtra)
library(dplyr)
load("~/GitHub/designedmissingness/Results/tables20180509_60_1_high_hightime0.RData")
high <- d

load("~/GitHub/designedmissingness/Results/tables20180509_60_1_med_hightime0.RData")
med <- d

load("~/GitHub/designedmissingness/Results/tables20180509_60_1_low_hightime0.RData")
low <- d

r <- rbind(low[[1]],med[[1]],high[[1]],low[[2]],med[[2]],high[[2]],low[[3]],med[[3]],high[[3]])

df <- data.frame(Design = rep(c("Split Form","Altered Split Form", "Wave Missingness"), each = 12), Missingness = rep(rep(c("Low","Medium", "High"), each = 4),3), Parameter = rep(c("$\\gamma_0$", "$\\beta_1$", "$\\beta_2$", "$\\beta_3$"),9), Bias = unlist(r["Bias"]), unlist(r["Pct Bias"]), MSE = unlist(r["MSE"]), Coverage = r["Coverage"], r["CI Length"], FMI = r["FMI"])

df[,4:9] <- apply(df[,4:9], 2, round, digits = 3)

names(df) <- c("Design", "Level", "Parameter", "Bias", "Pct Bias", "MSE", "Coverage", "CI Length", "FMI")
row.names(df) <- NULL
primary <- kable(df, escape = FALSE, format ="latex", caption = "Primary Results (N = 60)") %>%
  collapse_rows(columns = 1:2, latex_hline = "full") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position"))
@

<<sampsize, echo=FALSE, message=FALSE, warning=FALSE, fig.pos = "H">>=
library(kableExtra)
library(dplyr)
load("~/GitHub/designedmissingness/Results/tables20180509_120_1_high_hightime0.RData")
high <- d

load("~/GitHub/designedmissingness/Results/tables20180509_120_1_med_hightime0.RData")
med <- d

load("~/GitHub/designedmissingness/Results/tables20180509_120_1_low_hightime0.RData")
low <- d

r <- rbind(low[[1]],med[[1]],high[[1]],low[[2]],med[[2]],high[[2]],low[[3]],med[[3]],high[[3]])

df <- data.frame(Design = rep(c("Split Form","Altered Split Form", "Wave Missingness"), each = 12), Missingness = rep(rep(c("Low","Medium", "High"), each = 4),3), Parameter = rep(c("$\\gamma_0$", "$\\beta_1$", "$\\beta_2$", "$\\beta_3$"),9), Bias = unlist(r["Bias"]), unlist(r["Pct Bias"]), MSE = unlist(r["MSE"]), Coverage = r["Coverage"], r["CI Length"], FMI = r["FMI"])

df[,4:9] <- apply(df[,4:9], 2, round, digits = 3)

names(df) <- c("Design", "Level", "Parameter", "Bias", "Pct Bias", "MSE", "Coverage", "CI Length", "FMI")
row.names(df) <- NULL
sampsize <- kable(df, escape = FALSE, format ="latex", caption = "Increased Sample Size Results (N = 120)") %>%
  collapse_rows(columns = 1:2, latex_hline = "full") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position"))
@

<<hightime, echo=FALSE, message=FALSE, warning=FALSE, fig.pos = "H">>=
library(kableExtra)
library(dplyr)
load("~/GitHub/designedmissingness/Results/tables20180509_60_1_high_hightime1.RData")
high <- d

load("~/GitHub/designedmissingness/Results/tables20180509_60_1_med_hightime1.RData")
med <- d

load("~/GitHub/designedmissingness/Results/tables20180509_60_1_low_hightime1.RData")
low <- d

r <- rbind(low[[1]],med[[1]],high[[1]],low[[2]],med[[2]],high[[2]],low[[3]],med[[3]],high[[3]])

df <- data.frame(Design = rep(c("Split Form","Altered Split Form", "Wave Missingness"), each = 12), Missingness = rep(rep(c("Low","Medium", "High"), each = 4),3), Parameter = rep(c("$\\gamma_0$", "$\\beta_1$", "$\\beta_2$", "$\\beta_3$"),9), Bias = unlist(r["Bias"]), unlist(r["Pct Bias"]), MSE = unlist(r["MSE"]), Coverage = r["Coverage"], r["CI Length"], FMI = r["FMI"])

df[,4:9] <- apply(df[,4:9], 2, round, digits = 3)

names(df) <- c("Design", "Level", "Parameter", "Bias", "Pct Bias", "MSE", "Coverage", "CI Length", "FMI")
row.names(df) <- NULL
hightime <- kable(df, escape = FALSE, format ="latex", caption = "High Time Results (N = 60)") %>%
  collapse_rows(columns = 1:2, latex_hline = "full") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position"))
@

<<lowcorr, echo=FALSE, message=FALSE, warning=FALSE, fig.pos = "H">>=
library(kableExtra)
library(dplyr)
load("~/GitHub/designedmissingness/Results/tables20180509_60_0.5_high_hightime0.RData")
high <- d

load("~/GitHub/designedmissingness/Results/tables20180509_60_0.5_med_hightime0.RData")
med <- d

load("~/GitHub/designedmissingness/Results/tables20180509_60_0.5_low_hightime0.RData")
low <- d

r <- rbind(low[[1]],med[[1]],high[[1]],low[[2]],med[[2]],high[[2]],low[[3]],med[[3]],high[[3]])

df <- data.frame(Design = rep(c("Split Form","Altered Split Form", "Wave Missingness"), each = 12), Missingness = rep(rep(c("Low","Medium", "High"), each = 4),3), Parameter = rep(c("$\\gamma_0$", "$\\beta_1$", "$\\beta_2$", "$\\beta_3$"),9), Bias = unlist(r["Bias"]), unlist(r["Pct Bias"]), MSE = unlist(r["MSE"]), Coverage = r["Coverage"], r["CI Length"], FMI = r["FMI"])

df[,4:9] <- apply(df[,4:9], 2, round, digits = 3)

names(df) <- c("Design", "Level", "Parameter", "Bias", "Pct Bias", "MSE", "Coverage", "CI Length", "FMI")
row.names(df) <- NULL
lowcorr <- kable(df, escape = FALSE, format ="latex", caption = "Low Inter-Survey Correlation Results (N = 60)") %>%
  collapse_rows(columns = 1:2, latex_hline = "full") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position"))
@

Table \ref{tab:primary}

<<echo = FALSE, message = FALSE, warning = FALSE, warning=FALSE>>=
primary
@

Table \ref{tab:sampsize}

<<echo = FALSE, message = FALSE, warning = FALSE, warning=FALSE>>=
sampsize
@

Table \ref{tab:hightime}

<<echo = FALSE, message = FALSE, warning = FALSE, warning=FALSE>>=
hightime
@

Table \ref{tab:lowcorr}

<<echo = FALSE, message = FALSE, warning = FALSE, warning=FALSE>>=
lowcorr
@

\end{document}